SPARQL on the Open, Decentralised Web

### 2016-08-03 dissertation/introduction.tex
Some fixes #17

Conceptually, an RDF statement is a triple, where the terms are a
subject, a predicate and an object. The building blocks for these
triples are literals, blank nodes and Internationalized Resource
Identifiers (IRI). IRIs builds on the familiar URL: Uniform Resource
Locator, informally known as a Web address,
e.g. \url{https://www.w3.org/RDF/} is an address for a document about
RDF. It is also the document that lists the specifications that the
following outline relies. 
On the Semantic Web, the URL is first generalised to a Uniform
Resource Identifier, URI, that can not only name a document on the
Web, but any resource. A resource can be a physical thing, like a
person, or an abstract thing, like the type of weather at some place
at some time, numbers, strings, etc. Then, URIs are extended to IRIs
to enable the use of any character in them. Literals are strings with
a datatype (denoted by a IRI), and optionally a language tag. Finally,
blank nodes are unnamed nodes. RDF defines which of these building
blocks may be used for what term: The subject may be a IRI or a blank
node, the predicate only a IRI, whereas the object may be a literal,
blank node or IRI.

The use of IRIs makes it possible to view RDF statements as a directed
graph, where the subject IRI is a vertex, the predicate is an edge to
an object vertex\footnote{A working draft of the RDF Semantics
  specification dated 23 January 2003, see
  \url{https://www.w3.org/TR/2003/WD-rdf-mt-20030123/}, described the
  graph structure as a partially labeled directed pseudograph with
  unique node labels.}.

@@

SPARQL also introduced a mechanism for naming a graph or parts of a
graph with a IRI, an approach that was subsequently adopted by
RDF. Therefore, RDF is today often implemented not just a expressed as
a triple, but may be a expressed with a quad.

>>>>

Conceptually, an RDF statement is a triple, where the terms are a
subject, a predicate and an object. The building blocks for these
triples are literals, blank nodes and Internationalized Resource
Identifiers (IRI). IRIs builds on the familiar URL: Uniform Resource
Locator, informally known as a Web address,
e.g. \url{https://www.w3.org/RDF/} is an address for a document about
RDF. It is also the document that lists the specifications that the
following outline relies on. 
On the Semantic Web, the URL is first generalised to a Uniform
Resource Identifier, URI, that can not only name a document on the
Web, but any resource. A resource can be a physical thing, like a
person, or an abstract thing, like the type of weather at some place
at some time, numbers, strings, etc. Then, URIs are extended to IRIs
to enable the use of any character in them. Literals are strings with
a datatype (denoted by a IRI), and optionally a language tag. Finally,
blank nodes are unnamed nodes. RDF defines which of these building
blocks may be used for what term: The subject may be a IRI or a blank
node, the predicate only a IRI, whereas the object may be a literal,
blank node or IRI.

The use of IRIs makes it possible to view RDF statements as a directed
graph, where the subject IRI is a vertex, the predicate is an edge to
an object vertex\footnote{A working draft of the RDF Semantics
  specification dated 23 January 2003, see
  \url{https://www.w3.org/TR/2003/WD-rdf-mt-20030123/}, described the
  graph structure as a ``partially labeled directed pseudograph with
  unique node labels''.}.

@@

SPARQL also introduced a mechanism for naming a graph or parts of a
graph with a IRI, an approach that was subsequently adopted by
RDF. Therefore, RDF is today often implemented not just expressed as
a triple, but may be a expressed with a quad.
